---
title: "Day 6"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Challenge 1

We need to find the total number of questions that had at least one yes in each group, then add them. We can combine all answers in each group, then run `unique` |> `length` by group to find the total number of questions with at least one yes.  

```{r}
library(tidyr)
library(readr)
library(stringr)
library(purrr)

read_file(here::here("data/input_6.txt")) %>% 
  str_split("\n\n") %>% unlist() -> customs

customs %>% 
  str_replace_all("\n", "") %>% 
  str_split("") %>% 
  map_int(~length(unique(.))) %>% 
  sum()
```

## Challenge 2

Now we need the number of questions where everyone answered yes. The code is relatively straightforward except the `setdiff` call, which is there to remove the blank row at the end of the file. `Reduce(intersect)` finds the intersect of all of the vectors in the list. 

```{r}
customs %>% 
  str_split("\n") %>% 
  map_int(. %>% 
        setdiff("") %>% 
        str_split("") %>% 
        Reduce(intersect, .) %>% 
        length()
      ) %>% 
  sum()
```

