---
title: "Day 1"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Challenge 1

Find the two numbers in the input file that sum to 2020, return their product

```{r}
library(readr)

input <- as.numeric(read_lines(here::here("data/input_1_1.txt"))) 

two_numbers <- function(numbers, target){
  prod(intersect(numbers, target - numbers))
}

two_numbers(input, 2020)
```

## Challenge 2

Find the *three* numbers that sum to 2020 this time. My first thought was to apply the same  kind of logic as the two number solution, but in a loop with varying targets. 

```{r}
three_numbers_ugly <- function(numbers, target){
  temp <- target - numbers
  for(i in 1:length(numbers)){
    if(length(intersect(numbers, temp - numbers[i])) > 0){
      return(prod(c(intersect(numbers, temp - numbers[i]), numbers[i])))
    }
  }
}

three_numbers_ugly(input, 2020)
```

Since this is really the same logic, there's a much cleaner-looking version of this code that calls the two number function:

```{r}
three_numbers_cleaner <- function(numbers, target){
  targets <- target - numbers
  for(i in 1:length(targets)){
    if(two_numbers(numbers, targets[i]) > 1){
      return(two_numbers(numbers, targets[i]) * numbers[i])
    }
  }
}

three_numbers_cleaner(input, 2020)
```

People always say for loops are bad in R, so lets try a version without a loop using `purrr::map`.

```{r}
library(purrr)
library(dplyr)

three_numbers <- function(numbers, target){
  map_dbl(target - numbers, ~two_numbers(numbers, .)) %>% 
    prod() %>% 
    sqrt()
}

three_numbers(input, 2020)
```

Which is the fastest? 

```{r}
library(rbenchmark)

benchmark("ugly" = {three_numbers_ugly(input, 2020)},
          "cleaner" = {three_numbers_cleaner(input, 2020)},
          "map" = {three_numbers(input, 2020)}) %>% 
  knitr::kable()
```

The function using `purrr::map` is slower by an order of magnitude, partly because it doesn't stop early if the solution is found. 
