---
title: "Day 2"
output: md_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Challenge 1

```{r}
library(readr)

input <- read_lines(here::here("data/input_2.txt"))
```

This is a string processing problem. Given a string of the form "X-Y C: ccccc", determine whether C appears between X and Y times in c. The first step will be to parse the input strings into individual variables. `stringr::str_split()` should be useful for this.

```{r}
library(stringr)

str_split("X-Y C: ccccc", " |-|: ")
```

Next, we need a way to count the number of times our third variable appears in the fourth - `stringr::str_count()`.

```{r}
str_count("abcde", "a")
```

Putting it together:

```{r}
library(dplyr)

password_valid <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.numeric)) %>% 
    mutate(valid = between(str_count(str, char), low, high)) %>% 
    summarise(total_valid = sum(valid))
}

password_valid(input)
```

But this function doesn't get the right answer. Let's do some digging.

```{r}
test_case <- c("1-3 a: abcde",
               "1-3 b: cdefg",
               "2-9 c: ccccccccc")

password_valid <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.numeric)) %>% 
    mutate(valid = between(str_count(str, char), low, high)) #%>% 
    #summarise(total_valid = sum(valid))
}

password_valid(test_case)
```
Looks like there's something wrong with the validity logic. Test case 3 should be TRUE, not false. 

```{r}
password_valid <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.numeric)) %>% 
    mutate(count = str_count(str, char)) %>%
    mutate(valid = between(count, low, high))
    #summarise(total_valid = sum(valid))
}

password_valid(test_case)
```

Ugh. Floating point comparisons. Easy fix:


```{r}
library(dplyr)

password_valid <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.integer)) %>% # as.integer instead of as.numeric
    mutate(valid = between(str_count(str, char), low, high)) #%>% 
    #summarise(total_valid = sum(valid))
}

password_valid(test_case)
```

WTF?

```{r}
password_valid <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.integer)) %>% 
    mutate(val_count = str_count(str, char)) %>% 
    mutate(valid = val_count >= low & val_count <= high) %>% 
    summarise(total_valid = sum(valid))
}

password_valid(test_case)
```

Seems like the problem is the `dplyr::between()` function. Weird. 

```{r}
password_valid <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.numeric)) %>% 
    mutate(char_count = str_count(str, char)) %>%
    mutate(valid = char_count >= low & char_count <= high) %>% 
    summarise(total_valid = sum(valid))
}

password_valid(input)
```

## Challenge 2

The numbers actually describe positions in the password, not counts. Password is valid if either numbered position contains the character. We can modify the validity checks using the base `substr()` function to get characters at the required positions. Since exactly one of the positions is required to have the character, we need to combine the two comparisons using `xor()` rather than the typical `|`. 

```{r}
password_valid_positions <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.numeric)) %>% 
    mutate(char_1 = substr(str, low, low), char_2 = substr(str, high, high)) %>%
    mutate(valid = xor(char == char_1, char == char_2)) %>% 
    summarise(total_valid = sum(valid))
}

password_valid_positions(input)
```

## Wrap Up

We can clean up these solutions a little to avoid repeating code:

```{r}
library(readr)
library(stringr)
library(dplyr)

input <- read_lines(here::here("data/input_2.txt"))

extract <- function(string_vector){
  str_split(string_vector, " |-|: ") %>% 
    do.call(rbind.data.frame, .) %>% 
    magrittr::set_names(c("low", "high", "char", "str")) %>% 
    mutate(across(low:high, as.numeric))
}

valid_between <- function(dataframe){
  dataframe %>% 
    mutate(char_count = str_count(str, char)) %>%
    mutate(valid = char_count >= low & char_count <= high) %>% 
    summarise(total_valid = sum(valid))
}

valid_at <- function(dataframe){
  dataframe %>% 
    mutate(char_1 = substr(str, low, low), char_2 = substr(str, high, high)) %>%
    mutate(valid = xor(char == char_1, char == char_2)) %>% 
    summarise(total_valid = sum(valid))
}

extract(input) %>% valid_between()
extract(input) %>% valid_at()
```

