---
title: "Day 8"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Challenge 1

I think we can solve this one with a graph as well.

To build the graph from the input data:
  1. Every row is a node
  2. From each acc or nop row, there is an edge connecting to the next row
  3. Each jmp node has an edge connecting to the number indicated

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(igraph)

nodes <- read_lines(here::here("data/input_8.txt")) %>% 
  as_tibble() %>% 
  separate(value, into = c("command", "num"), sep = " ") %>%
  mutate(num = as.numeric(num)) %>% 
  add_row(command = "end", num = 0) %>% 
  tibble::rowid_to_column("id")

edges <- nodes %>% 
    rename(from = id) %>%
    mutate(to = if_else(command == "jmp", from + num, from + 1)) %>% 
    filter(to <= max(nodes$id)) %>% 
    mutate(weight = 0) %>% 
    select(from, to, weight)
```

```{r}
command_net <- graph_from_data_frame(vertices = nodes, d = edges, directed = TRUE)
```

```{r}
all_simple_paths(command_net, from = 1, mode = "out") %>% 
  last() %>% as.vector() %>% 
  as_tibble() %>% 
  left_join(nodes, by = c("value" = "id")) %>% 
  filter(command == "acc") %>% 
  summarize(total = sum(num)) %>% 
  pull()
```

## Challenge 2

Here we build a graph that includes all possible changes, but gives the changed edges a weight of 1. Finding the shortest path will find the path with the fewest changes (in this case, 1):

```{r}
edges_2 <- nodes %>% 
    rename(from = id) %>%
    mutate(to = if_else(command == "nop", from + num, Inf)) %>% 
    mutate(to = if_else(command == "jmp", from + 1, Inf)) %>% 
    filter(to <= max(nodes$id)) %>% 
    mutate(weight = 1) %>% 
    select(from, to, weight) %>% 
    rbind(edges)

all_possible <- graph_from_data_frame(d = edges_2, vertices = nodes, directed = TRUE)
```

```{r}
shortest_paths(all_possible, from = 1, to = 613)$vpath %>% 
  unlist() %>% as.vector() %>% as_tibble() %>% 
  left_join(nodes, by = c("value" = "id")) %>% 
  filter(command == "acc") %>% 
  summarize(total = sum(num)) %>% 
  pull()
```

