---
title: "Day 3"
output: md_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Challenge 1

We're given a text file consisting of . and # marks representing a map. We should be able to use `readr::read_fwf()` to turn this into a matrix:

```{r}
library(readr)

input <- read_fwf(here::here("data/input_3.txt"), fwf_widths(rep(1, times = 31)))
```

The challenge asks us to count trees starting from the top left and progressing right 3, down 1 each time. The landscape repeats to the right. Instead of actually repeating the landscape, we can "wrap" the horizontal index using a modulus.

```{r}
library(purrr)

map_dbl(1:15, ~1 + ((3 * (. - 1)) %% 31))
```

```{r}
trees_hit <- function(landscape, slope){
  h_ind <- map_dbl(1:nrow(landscape), ~1 + ((slope * (. - 1)) %% ncol(landscape)))
  sum(map2_lgl(1:nrow(landscape), h_ind, ~landscape[.x, .y] == "#"))
}

trees_hit(input, 3)
```

# Challenge 2

The function as written will work for part 2, except for the last slope (Right 1, Down 2).

```{r}
map_dbl(1:15, ~1 + ((.5 * .) %% 31))
```

Slopes that are more "downhill" than right 1, down 1 produce fractional indexes in the current function. We can just filter the fractional indexes out (note the use of the "explode" pipe `%$%`):

```{r}
library(dplyr)
library(magrittr)

trees_hit <- function(landscape, slope){
  data.frame(v = 1:nrow(landscape)) %>% 
    mutate(h = map_dbl(v, ~1 + ((slope * (. - 1)) %% ncol(landscape)))) %>%
    filter(h %% 1 == 0) %$%   
    map2_lgl(v, h, ~landscape[.x, .y] == "#") %>% 
    sum()
}

map_dbl(c(0.5, 1, 3, 5, 7), ~trees_hit(input, .)) %>% prod()
```


